---
title: "AP Assignment"
author: 'Poonam Nair [sid: 15136653]'
date: '2022-06-05'
output: word_document
---

By including this statement, we the authors of this work, verify that:
* We hold a copy of this assignment that we can produce if the original is lost or damaged.
* We hereby certify that no part of this assignment/product has been copied from any other student’s
work or from any other source except where due acknowledgement is made in the assignment.
No part of this assignment/product has been wrien/produced for us by another person except
where such collaboration has been authorised by the subject lecturer/tutor concerned.
* We are aware that this work may be reproduced and submied to plagiarism detection soware programs for the purpose of detecting possible plagiarism (which may retain a copy on its database
for future plagiarism checking).
* We hereby certify that we have read and understand what the School of Computing, Engineering
and Mathematics denes as minor and substantial breaches of misconduct as outlined in the learning
guide for this unit.

# QUESTION 1
## Write the code to compute the total scored by each team for each match. Note that each goal is worth 10 points and the team that caught the snitch is awarded an extra 50 points. Run your code on the data and present the results in a table showing the season and round number, the match number (a new variable that identies each match of the season), the home and away team names and their scores.

## DATA PREPROCESSING

```{r}
df <- read.csv("competitionResults.csv")

head(df)
```

## THE ALGORITHM

```{r}
# [RESULT] Total scored by each team for each match. Each goal is worth 10 points, team that caught the snitch  awarded extra 50 points. 
# * the season and round number
# * the match number (a new variable that identifies each match of the season)
# * the home and away team names and their scores
seasonTable <- data.frame(matrix(ncol=7, nrow=0))
headings <- c("season", "round", "matchNo", "homeTeam", "homeTeamPoints", "awayTeam", "awayTeamPoints")
colnames(seasonTable) <- headings
print(seasonTable)


for (match_number in 1:nrow(df)) {
  match <- df[match_number,]
  totalPointsHomeTeam <- match[["homeGoals"]] * 10 + ifelse(match[["homeTeam"]] == match[["snitch"]], 50, 0)
  totalPointsAwayTeam <- match[["awayGoals"]] * 10 + ifelse(match[["awayTeam"]] == match[["snitch"]], 50, 0)

  seasonTable[match_number, 1] <- match[["Season"]]
  seasonTable[match_number, 2] <- match[["round"]]
  seasonTable[match_number, 3] <- match_number
  seasonTable[match_number, 4] <- match[["homeTeam"]]
  seasonTable[match_number, 5] <- totalPointsHomeTeam
  seasonTable[match_number, 6] <- match[["awayTeam"]]
  seasonTable[match_number, 7] <- totalPointsAwayTeam
}

head(seasonTable)

```

# QUESTION 2
## The winner of each match has highest score. Each team scores 3 points for a win, 1 point for a draw. Team with most points at the end of season is the season winner. 

```{r}
teams_associatedPoints <- data.frame(matrix(ncol=2, nrow=0))
headings <- c("teamName", "totalPoints")
colnames(teams_associatedPoints) <- headings

# I built a function and passed a set (dataframe) and n (first n in set) variable
buildResultsLadder <- function(set, n_matches) {
  # aggregate points for each unique team in set
  for(i in 1:n_matches) {
    match <- set[i,]
    
    # award points: win, draw
    if(match[["homeTeamPoints"]] > match[["awayTeamPoints"]]) {
      homeTeamFinalPoints <- match[["homeTeamPoints"]] + 3
      awayTeamFinalPoints <- match[["awayTeamPoints"]]
    } else if(match[["homeTeamPoints"]] == match[["awayTeamPoints"]]) {
      homeTeamFinalPoints <- match[["homeTeamPoints"]] + 1
      awayTeamFinalPoints <- match[["awayTeamPoints"]] + 1
    } else if(match[["homeTeamPoints"]] < match[["awayTeamPoints"]]){
      homeTeamFinalPoints <- match[["homeTeamPoints"]]
      awayTeamFinalPoints <- match[["awayTeamPoints"]] + 3
    } else {
      # do nothing
    }
    
    # aggregate team points
    homeTeamIdx <- which(teams_associatedPoints$teamName == match[["homeTeam"]])
    awayTeamIdx <- which(teams_associatedPoints$teamName == match[["awayTeam"]])
    
    if(length(homeTeamIdx) != 0) {
      teams_associatedPoints[homeTeamIdx, 1] <- match[["homeTeam"]]
      teams_associatedPoints[homeTeamIdx, 2] <- homeTeamFinalPoints
    } else {
      # get next row index
      nextRowIdx <- ifelse(length(teams_associatedPoints$teamName) == 0, 1, length(teams_associatedPoints$teamName) + 1)
      teams_associatedPoints[nextRowIdx, 1] <- match[["homeTeam"]]
      teams_associatedPoints[nextRowIdx, 2] <- homeTeamFinalPoints
    }
    
    if(length(awayTeamIdx) != 0) {
      teams_associatedPoints[awayTeamIdx, 1] <- match[["awayTeam"]]
      teams_associatedPoints[awayTeamIdx, 2] <- awayTeamFinalPoints
    } else {
      nextRowIdx <- ifelse(length(teams_associatedPoints$teamName) == 0, 1, length(teams_associatedPoints$teamName) + 1)
      teams_associatedPoints[nextRowIdx, 1] <- match[["awayTeam"]]
      teams_associatedPoints[nextRowIdx, 2] <- awayTeamFinalPoints
    }
    
  }
  # Order teams by points descending
  teams_associatedPoints_ordered <- order(teams_associatedPoints$totalPoints, decreasing = TRUE) 
  # returns index of order, used to ORDER tables
  resultsLadder <- teams_associatedPoints[teams_associatedPoints_ordered,]
  print(resultsLadder)
}

# [RESULT] Results Ladder: Each team’s points as per the first 5 rounds (ordered by points)
first5MatchesSet <- seasonTable[1:5,]
first_n_matches <- length(first5MatchesSet$season)
buildResultsLadder(first5MatchesSet, first_n_matches)

# [RESULT] Results Ladder: Each team’s points for the full season (ordered by points)
season2016Set <- subset(seasonTable, subset=(season == 2016))
season_no_matches <- length(season2016Set$season)
buildResultsLadder(season2016Set, season_no_matches)

# Pick the season who’s last digit corresponds to 3.
season2013Set <- subset(seasonTable, subset=(season == 2013))
season_no_matches <- length(season2013Set$season)
buildResultsLadder(season2013Set, season_no_matches)

```

# QUESTION 3
## An investigation into home game advantage is planned for the coming months. Write a function that returns a count of the number of home games each team has had, the total number of games each team has played in, the total points scored in a home game and the total number of points the team has scored. Provide these results in a data frame. Run the function on the provided data and show the resulting table output. 

Function below with associated `getTotalPoints()` I've created to handle point/goal calculations. Scroll to next code snippet to test. code

```{r}
## Please note: normally, I would use more of a functional programming approach (centralisation and usage of modules), but for time efficiency, I have not fully done this. 
homeGameCalculations <- function (cf) {
  # numeric vars
  n_matches = nrow(competition.results.df)
  # data vars
  team.statistics <- data.frame(matrix(ncol=5, nrow=0))
  colnames(team.statistics) <- c("team", "home.games", "total.games", "total.points.homegame", "total.points")

  for(i in 1:n_matches) {
    match <- competition.results.df[i,]
    homeTeam <- match$homeTeam
    awayTeam <- match$awayTeam
  
    # indexes
    homeTeam.idx <- which(team.statistics$team == homeTeam)
    awayTeam.idx <- which(team.statistics$team == awayTeam)
    nextIdx <- length(team.statistics$team)
  
    # Hometeam stat calculations
    homeIdx <- ifelse(nextIdx == 0, 1, ifelse(length(homeTeam.idx) != 0, homeTeam.idx, nextIdx+1))
    
    # Awayteam stat calculations
    awayIdx <- ifelse(nextIdx == 0, 2, ifelse(length(awayTeam.idx) != 0, awayTeam.idx, nextIdx+2))
  
    team.statistics[homeIdx, 1] <- homeTeam 
    team.statistics[awayIdx, 1] <- awayTeam 
    
    # The number of HOME games each team has had
    totalHomeGames <- ifelse(is.na(team.statistics[homeIdx, 2]), 1, team.statistics[homeIdx, 2] + 1)
    team.statistics[homeIdx, 2] <- totalHomeGames 
    team.statistics[awayIdx, 2] <- ifelse(is.na(team.statistics[awayIdx, 2]), 0, team.statistics[awayIdx, 2])
  
    # The total number of games each team has played in
    totalHomeTeamGames <- ifelse(is.na(team.statistics[homeIdx, 3]), 1, team.statistics[homeIdx, 3] + 1)
    team.statistics[homeIdx, 3] <- totalHomeTeamGames 
    totalAwayTeamGames <- ifelse(is.na(team.statistics[awayIdx, 3]), 1, team.statistics[awayIdx, 3] + 1)
    team.statistics[awayIdx, 3] <- totalAwayTeamGames 
  
    # The total points scored in a home game
    totalHomeGamePoints <- ifelse(is.na(team.statistics[homeIdx, 4]), 1, team.statistics[homeIdx, 4] + match$homeGoals)
    team.statistics[homeIdx, 4] <- totalHomeGamePoints
    team.statistics[awayIdx, 4] <- ifelse(is.na(team.statistics[awayIdx, 4]), 0, team.statistics[awayIdx, 4])
  
    # total points overall
    totalPointsForMatch <- 0 # initialise variable here
    whoGotSnitch <- match$snitch
        totalPointsForMatch <-  getTotalPoints(homeTeam, awayTeam, match$homeGoals, match$awayGoals, totalHomeGamePoints, team.statistics[awayIdx, 5], whoGotSnitch)
  
    team.statistics[homeIdx, 5] <- totalPointsForMatch[1]
    team.statistics[awayIdx, 5] <- totalPointsForMatch[2]
  }
  
  print(team.statistics)

}

getTotalPoints <- function(homeTeam, awayTeam, homeTeamGoals, awayTeamGoals, homeTeamPoints, awayTeamPoints, whoGotSnitch) {
  homeTeamFinalPoints <- 0
  awayTeamFinalPoints <- 0
  
  # convert goals to points
  homeTeamFinalPoints <- homeTeamGoals * 10
  awayTeamFinalPoints <- awayTeamGoals * 10

  # check NA on awayTeamPoints
  awayTeamPoints <- ifelse(is.na(awayTeamPoints),0,awayTeamPoints)
  
  # award points: win, draw
  if(homeTeamPoints > awayTeamPoints) {
    homeTeamFinalPoints <- homeTeamPoints + 3
    awayTeamFinalPoints <- awayTeamPoints
  } else if(homeTeamPoints == awayTeamPoints) {
    homeTeamFinalPoints <- homeTeamPoints + 1
    awayTeamFinalPoints <- awayTeamPoints + 1
  } else if(homeTeamPoints < awayTeamPoints){
    homeTeamFinalPoints <- homeTeamPoints
    awayTeamFinalPoints <- awayTeamPoints + 3
  } else {
    # do nothing
  }
  
  # snitch bonus points
  if(homeTeam == whoGotSnitch) {
    homeTeamFinalPoints <- homeTeamFinalPoints + 50
  } else if (awayTeam == whoGotSnitch) {
    awayTeamFinalPoints <- awayTeamFinalPoints + 50
  }

  total.match.points <- c(homeTeamFinalPoints, awayTeamFinalPoints)
  total.match.points
}

competition.results.df <- read.csv("competitionResults.csv")
  
# Run the function on the provided data and show the resulting table output.
homeGameCalculations(competition.results.df)
```


## Testing function

```{r message=TRUE, warning=TRUE}
# PUT HERE
```

# QUESTION 4

## Provide a table containing the team statistics for all seasons. Including, but not limited to the following: number of games, number of wins, winning percentage, points scored, points conceded, number of snitch catches, goals scored, number of tournament wins. Order the teams in terms of total points scored.
```{r}
library(expss)

s1 <- data.frame(matrix(nrow=2,ncol=2))
colnames(s1) <- c("test","test1")
s1[2,1] <- "a"
s1[2,2] <- "a_test1"

s_1 <- list("2012", s1)

s2 <- data.frame(matrix(nrow=2,ncol=2))
colnames(s1) <- c("test","test1")
s2[2,1] <- "b"
s2[2,2] <- "b_test1"


s_1 <- list("2012", s1)
s_2 <- list("2013", s2)
nnn <- list()

nnn[[1]] <- s_1
# print(length(nnn))
#print(nnn[[2]][2])

```


```{r}

# Provide a table containing the team statistics for all seasons. 
c_data <- read.csv("competitionResults.csv")
seasonList <- list()
n_matches <- length(c_data$Season)

  for(i in 1:n_matches) {
    match <- c_data[i,]
    homeTeam <- match$homeTeam
    awayTeam <- match$awayTeam
    season <- match$Season
  
    ###############################
    # Season data
    ###############################
    # if seasonList does't have season, create, else extract subset
    seasonSubset <- data.frame(matrix(nrow=0,ncol=9))
    colnames(seasonSubset) <- c("team","no.games","no.wins","win.percentage","points.scored","points.conceded","total.snitches","goals","tournament.wins")
    seasonExistsIndex <- 0
    homeTeamExistsIndex <- 0
    awayTeamExistsIndex <- 0

    if(length(seasonList) > 0) {
      for(s in length(seasonList)) {
        if(season == seasonList[[s]][1]) {
          seasonSubset <- seasonList[[s]][2]
          # get match from season subset
          for(ss in 1:length(seasonSubset$no.games)) {
            if(seasonSubset$team == match$homeTeam) {
              homeTeam <- seasonSubset
              homeTeamExistsIndex <- ss
            }
            if(seasonSubset$team == match$awayTeam) {
              awayTeam <- seasonSubset
              awayTeamExistsIndex <- ss
            }
          }
          seasonExistsIndex <- s
        }
      }
    }
    
    ###############################
    # Match statistics
    ###############################
    # calculate
    # Number of games
    totalHomeGames <- ifelse(homeTeamExistsIndex == 0, 1, seasonSubset[homeTeamExistsIndex, 2] + 1)
    totalAwayGames <- ifelse(awayTeamExistsIndex == 0, 1, seasonSubset[awayTeamExistsIndex, 2] + 1)

    # Number of points
    totalHomePoints <- ifelse(homeTeamExistsIndex != 0, ifelse(is.na(seasonSubset[homeTeamExistsIndex, 5]),match$homeGoals * 10,seasonSubset[homeTeamExistsIndex, 5]), match$homeGoals * 10)
        totalAwayPoints <- ifelse(homeTeamExistsIndex != 0, ifelse(is.na(seasonSubset[homeTeamExistsIndex, 5]),match$homeGoals * 10,seasonSubset[homeTeamExistsIndex, 5]), match$homeGoals * 10)

    # Number of wins
    totalHomeWins <- ifelse(homeTeamExistsIndex == 0 && totalHomePoints > totalAwayPoints, seasonSubset[homeTeamExistsIndex, 3] + 1, seasonSubset[homeTeamExistsIndex, 3])
    totalAwayWins <- ifelse(awayTeamExistsIndex == 0 && totalHomePoints < totalAwayPoints, seasonSubset[awayTeamExistsIndex, 3] + 1, seasonSubset[awayTeamExistsIndex, 3])
    
    # Points conceded
    homePointsConceded <- ifelse(homeTeamExistsIndex != 0, ifelse(is.na(seasonSubset[homeTeamExistsIndex, 6]), totalAwayPoints, seasonSubset[homeTeamExistsIndex, 6] + totalAwayPoints), totalHomePoints)
    awayPointsConceded <- ifelse(awayTeamExistsIndex != 0, ifelse(is.na(seasonSubset[awayTeamExistsIndex, 6]), totalHomePoints, seasonSubset[awayTeamExistsIndex, 6] + totalHomePoints), totalAwayPoints)

                # colnames(seasonSubset) <- c("team","no.games","no.wins","win.percentage","points.scored","points.conceded","total.snitches","goals","tournament.wins")
    
    # Winning percentage
    totalHomeWinPerc <- totalHomeWins / totalHomeGames
    totalAwayWins <- totalAwayWins / totalAwayGames
    
    # no of snitch catches, goals scored, number of tournament wins.

    ###############################
    # Update/Insert season
    ###############################
    # insert partial data into season data
    
    h.team.idx = ifelse(homeTeamExistsIndex != 0, homeTeamExistsIndex, length(seasonSubset$team) + 1)
    seasonSubset[h.team.idx,1] <- homeTeam
    seasonSubset[h.team.idx,2] <- totalHomeGames
    seasonSubset[h.team.idx,3] <- totalHomeWins
    seasonSubset[h.team.idx,4] <- 0
    seasonSubset[h.team.idx,5] <- totalHomePoints
    seasonSubset[h.team.idx,6] <- homePointsConceded
    seasonSubset[h.team.idx,7] <- 0
    seasonSubset[h.team.idx,8] <- 0
    seasonSubset[h.team.idx,9] <- 0
      
    a.team.idx = ifelse(awayTeamExistsIndex != 0, awayTeamExistsIndex, length(seasonSubset$team) + 1)
    seasonSubset[a.team.idx,1] <- awayTeam
    seasonSubset[a.team.idx,2] <- totalAwayGames
    seasonSubset[a.team.idx,3] <- totalAwayWins
    seasonSubset[a.team.idx,4] <- 0
    seasonSubset[a.team.idx,5] <- totalAwayPoints
    seasonSubset[a.team.idx,6] <- awayPointsConceded
    seasonSubset[a.team.idx,7] <- 0
    seasonSubset[a.team.idx,8] <- 0
    seasonSubset[a.team.idx,9] <- 0

    # insert season subset back into season
    if(seasonExistsIndex != 0) {
      seasonList[[seasonExistsIndex]][[1]] <- season
      seasonList[[seasonExistsIndex]][[2]] <- seasonSubset
    }else {
      seasonList[[(length(seasonList) + 1)]] <- list(season, seasonSubset)
    }
  }

# TODO: I couldn't easily add subheaders for seasons, so I'm outputting the entire list instead
  # for(i in length(seasonList)) {
  #   print(seasonList[[i]][2][[1]])
  # }
  
  print(seasonList)
# TODO: Teams ordered in terms of total points scored.

```






